Evolution, functional diversity, regime shifts, hysteresis, and nonlinearity; ecosystem responses to environmental change.

### 2021-06-09 manuscript/article.Rmd
many additions

---
title: Evolution, functional diversity, regime shifts, hysteresis, and nonlinearity; ecosystem responses to environmental change.
subtitle: Predictions of a model of an oxic-anoxic microbial ecosystem 
authors:
  - name: Owen L Petchey
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: owen.petchey@ieu.uzh.ch
  - name: Rainer Krug
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: rainer.krug@ieu.uzh.ch
  - name: Marcel Suleiman
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: marcel.suleiman@ieu.uzh.ch

@@

```{r include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

@@

This report concerns how biodiversity influences the strength and nature of ecosystem responses to environmental change (figure 1: panel a shows linear response with different gradients, large being with low diversity; panel b shows a linear and a non-linear response starting and ending at the same place; panel c shows a hysteretic response, with greater or less range of hysteresis). Strong responses of ecosystems to environmental change imply high *sensitivity* and lack of resistance to environmental change, and higher temporal variability in the face of varying environmental conditions. Greater *non-linearity* in ecosystem responses increases the potential for surprises, such as when an ecosystem that was resistant to an environmental change becomes unexpectedly sensitive to further change. The presence and extent of *hysteresis* in ecosystem responses to environmental driver gradients concerns how readily are ecosystem responses reversed when environmental change is reversed.

In this context, a directly relevant aspect of biodiversity is variation in functional traits, i.e. the traits that determine organismal responses to environmental conditions and that determine the effects of organisms on their environment (citation). Variation in functional traits is often termed "functional diversity". There is a considerable amount of theory predicting, and evidence supporting, that greater functional diversity leads to lower temporal variability in ecosystem states variables (such as total biomass) (citations) (Leary, Pennekamp). The mechanism by which this occurs is that greater functional response diversity causes a weaker relationship between an environmental driver gradient and the ecosystem state (figure 1a), due to the functional turnover along the environmental driver gradient that results from variation in organisms' responses to the environmental driver gradient .

There is less theory about effects of functional diversity on the linearity (figure 1b) and the hysteresis (figure 1c) of ecosystem responses to environmental change. @dakos2019b, in their article *Ecosystem tipping points in an evolving world* asked how changes in functional trait distributions (including evolutionary changes) affect the occurrence of tipping points. They explain that:

@@

-   On the other hand, the opposite could occur: greater functional diversity could make a tipping point occur with less change in an environmental driver (i.e. may cause lower resilience) (Figure 2c in \@dakos2019b). This could occur if there is selection for greater tolerance is causes, incidentally, selection for lower growth rate, perhaps due to a tradeoff between the two traits.

@@

These explanations/predictions were based on the verbal/graphical argumentation @dakos2019b. There were no mathematical models of ecosystem responses to environmental change behind the explanations/predictions. @ceulemans2021, in their article *Functional diversity alters the effects of a pulse perturbation on the dynamics of tritrophic food webs*, did employ a such a mathematical model.

@@

## Scope of our investigation

Questions: - How does variation organismal tolerances affect response

Approach: - Modelling the system with ODEs. - Diversity manipulation (can be considered selection among strains, or selection among species). - Do not include mutation.

According the the questions (in the Introduction) this is what we did:

Variation in amount of among strain variation, no variation in number of strains, no variation in strength or shape of the of the tradeoff

Effect of among strain variation on how the steady state(s) of the system varies with in different environmental conditions, and on the dynamical response of system to gradual environmental change.

@@

@@

```{r tradeoff, out.width="0.5\\linewidth", fig.align="center", fig.cap=c("An example of among-strain variation in growth rate and tolerance, and the tradeoff in the variation."), echo = FALSE}
ggplot() +
  geom_point(aes(x = G_max_i, y = h_i)) +
  xlab("Maximum growth rate") +
  ylab("Tolerance") +
  theme_bw(base_size = 20)
```

@@

>>>>

---
title: Evolution, functional diversity, regime shifts, hysteresis, and nonlinearity; ecosystem responses to environmental change.
subtitle: Predictions of a model of an oxic-anoxic microbial ecosystem 
authors:
  - name: Owen L Petchey
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: owen.petchey@ieu.uzh.ch
  - name: Rainer Krug
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: rainer.krug@ieu.uzh.ch
  - name: Any Others
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: any.others@ieu.uzh.ch
  - name: Marcel Suleiman
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: marcel.suleiman@ieu.uzh.ch

@@

```{r include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE)
```

@@

This report concerns how biodiversity influences the strength and nature of ecosystem responses to environmental change (figure 1; panel(a) three ovals, environment oval with arrow to ecosystem property oval, with biodiversity oval pointing the arrow between environment and ecosystem property; panel (b) shows linear response with different gradients, large being with low diversity; panel (c) shows a linear and a non-linear response starting and ending at the same place; panel (d) shows a hysteretic response, with greater or less range of hysteresis). Strong responses of ecosystems to environmental change imply high *sensitivity* and lack of resistance to environmental change, and higher temporal variability in the face of varying environmental conditions. Greater *non-linearity* in ecosystem responses increases the potential for surprises, such as when an ecosystem that was resistant to an environmental change becomes unexpectedly sensitive to further change. The presence and extent of *hysteresis* in ecosystem responses to environmental driver gradients concerns how readily are ecosystem responses reversed when environmental change is reversed.

In this context, a directly relevant aspect of biodiversity is variation in functional traits, i.e. the traits that determine organismal responses to environmental conditions and that determine the effects of organisms on their environment (citation). Variation in functional traits is often termed "functional diversity". There is a considerable amount of theory predicting, and evidence supporting, that greater functional diversity leads to lower temporal variability in ecosystem states variables (such as total biomass) (citations) (Leary, Pennekamp). The mechanism by which this occurs is that greater functional response diversity causes a weaker relationship between an environmental driver gradient and the ecosystem state (figure 1a/b), due to the functional turnover along the environmental driver gradient that results from variation in organisms' responses to the environmental driver gradient (citation).

There is less theory about effects of functional diversity on the linearity (figure 1c) and the hysteresis (figure 1d) of ecosystem responses to environmental change. @dakos2019b, in their article *Ecosystem tipping points in an evolving world* asked how changes in functional trait distributions (including evolutionary changes) affect the occurrence of tipping points. They explain that:

@@

-   On the other hand, the opposite could occur: greater functional diversity could make a tipping point occur with less change in an environmental driver (i.e. may cause lower resilience) (Figure 2c in \@dakos2019b). This could occur if there is selection for greater tolerance that causes, incidentally, selection for lower growth rate, perhaps due to a tradeoff between the two traits.

@@

These explanations/predictions were based on the verbal/graphical argumentation @dakos2019b. There were no mathematical models of ecosystem responses to environmental change behind the explanations/predictions. @ceulemans2021, in their article *Functional diversity alters the effects of a pulse perturbation on the dynamics of tritrophic food webs*, did employ a such a mathematical model [to be developed].

@@

## Aim and scope of our investigation

Aim:

* To provide an answer to the question of how trait variation affects the non-linearity, resilience and hysteresis of ecosystem response to an environmental change.

Scope:

* Mathematical modelling of a microbrial ecosystem. Trait diversity manipulation (can be considered selection among strains, or selection among species). (Do not include mutation or recomination).

* Variation in amount of among strain variation, no variation in number of strains, no variation in strength or shape of the of the tradeoff.

* Measurement of the effect on how the steady state(s) of the system varies with in different environmental conditions, and on the dynamical response of system to gradual environmental change.

@@

"Variation in Sulfide Tolerance of Photosystem II in Phylogenetically Diverse Cyanobacteria from Sulfidic Habitats", by Miller and Bebout (2004) [https://aem.asm.org/content/70/2/736.short]

@@

```{r tradeoff, include = TRUE, echo = FALSE, out.width="0.5\\linewidth", fig.align="center", fig.cap=c("An example of among-strain variation in growth rate and tolerance, and the tradeoff in the variation.")}
ggplot() +
  geom_point(aes(x = G_max_i, y = h_i)) +
  xlab("Maximum growth rate") +
  ylab("Tolerance") +
  theme_bw(base_size = 20)
```

@@

## Further relaxing the assumptions

Deterministic -> stochastic : www.pnas.org/cgi/doi/10.1073/pnas.1414708112

### 2021-06-09 manuscript/article.tex
many additions

\author{
    Owen L Petchey
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:owen.petchey@ieu.uzh.ch}{\nolinkurl{owen.petchey@ieu.uzh.ch}}} \\
   \And
    Rainer Krug
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:rainer.krug@ieu.uzh.ch}{\nolinkurl{rainer.krug@ieu.uzh.ch}}} \\
   \And
    Marcel Suleiman
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:marcel.suleiman@ieu.uzh.ch}{\nolinkurl{marcel.suleiman@ieu.uzh.ch}}} \\
  }

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

When thinking about ecosystem stability, a key question is how
biodiversity influences ecosystem responses to environmental change
(figure 1: panel a shows linear response with different gradients, large
being with low diversity; panel b shows a linear and a non-linear
response starting and ending at the same place; panel c shows a
hysteretic response, with greater or less range of hysteresis).

In this context, a directly relevant aspect of biodiversity is variation
in functional traits, i.e.~the traits that determine organismal
responses to environmental conditions and that determine the effects of
organisms on their environment (citation). Variation in functional
traits is often termed ``functional diversity.''

There is a considerable amount of theory predicting, and evidence
supporting, that greater functional diversity leads to lower temporal
variability in ecosystem states variables (such as total biomass)
(citations). The mechanism by which this occurs is that greater
functional response diversity causes a weaker relationship between an
environmental driver gradient and the ecosystem state (figure 1a), due
to the functional turnover along the environmental driver gradient that
results from variation in organisms' responses to the environmental
driver gradient .

There is less theory about effects of functional diversity on the
linearity (figure 1b) and the hysteresis (figure 1c) of ecosystem
responses to environmental change. Dakos et al. (2019), in their article
``Ecosystem tipping points in an evolving world'' asked how changes in
functional trait distributions (including evolutionary changes) affect
the occurrence of tipping points. They predict that:

\begin{itemize}
\item
  In general, greater functional response diversity may decrease the
  likelihood of a catastrophic ecosystem response (Figure 2a in Dakos et
  al. (2019)). Here, we interpret this prediction to be that greater
  functional response diversity decreases the non-linearity of ecosystem
  response to environmental driver change, including the extreme
  non-linearity that can occur when an ecosystem ``flips'' from one
  state to another one.
\item
  Greater functional trait diversity can make a tipping point occur only
  with greater change in an environmental driver (Figure 2a in Dakos et
  al. (2019)). I.e. functional trait diversity can increase
  resilience--the extent of environmental change required to push the
  ecosystem into another state (Holling (1973)). This could occur if
  greater functional diversity equates to the presence of organisms with
  greater tolerance to the environmental driver Dakos et al. (2019).
\item
  On the other hand, the opposite could occur: greater functional
  diversity could make a tipping point occur with less change in an
  environmental driver (i.e.~may cause lower resilience) (Figure 2c in
  @dakos2019b). This could occur if there is selection for greater
  tolerance is causes, incidentally, selection for lower growth rate,
  perhaps due to a tradeoff between the two traits.
\item
  Finally, the amount of functional diversity may affect the ``recovery
  trajectory,'' i.e.~the relationship between ecosystem state and the
  environmental driver when the driver change is reversed Dakos et al.
  (2019). Put another way, functional diversity may increase or decrease
  the amount of hysteresis, and may even result in a recovery with lower
  non-linearity / no tipping-point (Figure 3 in Dakos et al. (2019)).
\end{itemize}

One reason why these predictions are often indeterminate is that trait
distributions can be affected by, and effect, organismal abundances and
environmental conditions. These numerous resulting feedbacks make
predictions from verbal argumentation rather difficult. \textbf{Discuss
Ceulemans, R., Wojcik, L.A. \& Gaedke, U. (2021). Functional diversity
alters the effects of a pulse perturbation on the dynamics of tritrophic
food webs. \emph{bioRxiv}, 2021.03.22.436420.}

\hypertarget{scope-of-our-investigation}{%
\subsection{Scope of our
investigation}\label{scope-of-our-investigation}}

Questions: - How does variation organismal tolerances affect response

Approach: - Modelling the system with ODEs. - Diversity manipulation
(can be considered selection among strains, or selection among species).
- Do not include mutation.

According the the questions (in the Introduction) this is what we did:

Variation in amount of among strain variation, no variation in number of
strains, no variation in strength or shape of the of the tradeoff

Effect of among strain variation on how the steady state(s) of the
system varies with in different environmental conditions, and on the
dynamical response of system to gradual environmental change.

@@

@@

@@

@@

@@

Need a more process based approach to variation in biodiversity:
mutation, genetics,

As mentioned in the Introduction, the numerous feedacks among trait
distributions, organismal abundances, and environmental conditions,
resulting feedbacks make predictions from verbal argumentation rather
difficult. They may also make for context dependent effects of
functional diversity. I.e. functional diversity may delay a tipping
point in one context (e.g.~ecosystem type) but delay it in another. This
is why we need more investigations, both theoretical and empirical, of
effects of functional trait diversity on environmental responses to
environmental change.

@@

@@

>>>>

\author{
    Owen L Petchey
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:owen.petchey@ieu.uzh.ch}{\nolinkurl{owen.petchey@ieu.uzh.ch}}} \\
   \And
    Rainer Krug
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:rainer.krug@ieu.uzh.ch}{\nolinkurl{rainer.krug@ieu.uzh.ch}}} \\
   \And
    Any Others
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:any.others@ieu.uzh.ch}{\nolinkurl{any.others@ieu.uzh.ch}}} \\
   \And
    Marcel Suleiman
   \\
    Department of Evolutionary Biology and Environmental Studies \\
    University of Zurich \\
  Zurich 8057, Switzerland \\
  \texttt{\href{mailto:marcel.suleiman@ieu.uzh.ch}{\nolinkurl{marcel.suleiman@ieu.uzh.ch}}} \\
  }

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

This report concerns how biodiversity influences the strength and nature
of ecosystem responses to environmental change (figure 1; panel(a) three
ovals, environment oval with arrow to ecosystem property oval, with
biodiversity oval pointing the arrow between environment and ecosystem
property; panel (b) shows linear response with different gradients,
large being with low diversity; panel (c) shows a linear and a
non-linear response starting and ending at the same place; panel (d)
shows a hysteretic response, with greater or less range of hysteresis).
Strong responses of ecosystems to environmental change imply high
\emph{sensitivity} and lack of resistance to environmental change, and
higher temporal variability in the face of varying environmental
conditions. Greater \emph{non-linearity} in ecosystem responses
increases the potential for surprises, such as when an ecosystem that
was resistant to an environmental change becomes unexpectedly sensitive
to further change. The presence and extent of \emph{hysteresis} in
ecosystem responses to environmental driver gradients concerns how
readily are ecosystem responses reversed when environmental change is
reversed.

In this context, a directly relevant aspect of biodiversity is variation
in functional traits, i.e.~the traits that determine organismal
responses to environmental conditions and that determine the effects of
organisms on their environment (citation). Variation in functional
traits is often termed ``functional diversity.'' There is a considerable
amount of theory predicting, and evidence supporting, that greater
functional diversity leads to lower temporal variability in ecosystem
states variables (such as total biomass) (citations) (Leary, Pennekamp).
The mechanism by which this occurs is that greater functional response
diversity causes a weaker relationship between an environmental driver
gradient and the ecosystem state (figure 1a/b), due to the functional
turnover along the environmental driver gradient that results from
variation in organisms' responses to the environmental driver gradient
(citation).

There is less theory about effects of functional diversity on the
linearity (figure 1c) and the hysteresis (figure 1d) of ecosystem
responses to environmental change. Dakos et al. (2019), in their article
\emph{Ecosystem tipping points in an evolving world} asked how changes
in functional trait distributions (including evolutionary changes)
affect the occurrence of tipping points. They explain that:

\begin{itemize}
\item
  In general, greater functional response diversity may decrease the
  likelihood of a catastrophic ecosystem response (Figure 2a in Dakos et
  al. (2019)). Here, we interpret this prediction to be that greater
  functional response diversity decreases the non-linearity of ecosystem
  response to environmental driver change, including the extreme
  non-linearity that can occur when an ecosystem ``flips'' from one
  state to another one.
\item
  Greater functional trait diversity can make a tipping point occur only
  with greater change in an environmental driver (Figure 2a in Dakos et
  al. (2019)). I.e. functional trait diversity can increase
  resilience--the extent of environmental change required to push the
  ecosystem into another state (Holling (1973)). This could occur if
  greater functional diversity equates to the presence of organisms with
  greater tolerance to the environmental driver Dakos et al. (2019).
\item
  On the other hand, the opposite could occur: greater functional
  diversity could make a tipping point occur with less change in an
  environmental driver (i.e.~may cause lower resilience) (Figure 2c in
  @dakos2019b). This could occur if there is selection for greater
  tolerance that causes, incidentally, selection for lower growth rate,
  perhaps due to a tradeoff between the two traits.
\item
  Finally, the amount of functional diversity may affect the ``recovery
  trajectory,'' i.e.~the relationship between ecosystem state and the
  environmental driver when the driver change is reversed Dakos et al.
  (2019). Put another way, functional diversity may increase or decrease
  the amount of hysteresis, and may even result in a recovery with lower
  non-linearity / no tipping-point (Figure 3 in Dakos et al. (2019)).
\end{itemize}

These explanations/predictions were based on the verbal/graphical
argumentation Dakos et al. (2019). There were no mathematical models of
ecosystem responses to environmental change behind the
explanations/predictions. Ceulemans, Wojcik, and Gaedke (2021), in their
article \emph{Functional diversity alters the effects of a pulse
perturbation on the dynamics of tritrophic food webs}, did employ a such
a mathematical model {[}to be developed{]}.

We discuss empirical studies in the Discussion section (spoiler--there
are few, if any, directly relevant ones).

\hypertarget{aim-and-scope-of-our-investigation}{%
\subsection{Aim and scope of our
investigation}\label{aim-and-scope-of-our-investigation}}

Aim:

\begin{itemize}
\tightlist
\item
  To provide an answer to the question of how trait variation affects
  the non-linearity, resilience and hysteresis of ecosystem response to
  an environmental change.
\end{itemize}

Scope:

\begin{itemize}
\item
  Mathematical modelling of a microbrial ecosystem. Trait diversity
  manipulation (can be considered selection among strains, or selection
  among species). (Do not include mutation or recomination).
\item
  Variation in amount of among strain variation, no variation in number
  of strains, no variation in strength or shape of the of the tradeoff.
\item
  Measurement of the effect on how the steady state(s) of the system
  varies with in different environmental conditions, and on the
  dynamical response of system to gradual environmental change.
\end{itemize}

@@

``Variation in Sulfide Tolerance of Photosystem II in Phylogenetically
Diverse Cyanobacteria from Sulfidic Habitats,'' by Miller and Bebout
(2004) {[}\url{https://aem.asm.org/content/70/2/736.short}{]}

@@

\hypertarget{model-dimensions}{%
\subsection{Model dimensions}\label{model-dimensions}}

\begin{itemize}
\tightlist
\item
  Time: hours
\item
  Volume: litres
\item
  Substrate quantity: micromoles
\item
  Organism quantity: cells
\end{itemize}

@@

{\centering \includegraphics[width=0.5\linewidth]{article_files/figure-latex/tradeoff-1}

\caption{An example of among-strain variation in growth rate and tolerance, and the tradeoff in the variation.}\label{fig:tradeoff}
\end{figure}

@@

One reason why the explanations of Dakos et al are often indeterminate
is that trait distributions can be affected by, and effect, organismal
abundances and environmental conditions. These numerous resulting
feedbacks may limit the determinancy of explanations based on the
verbal/graphical argumentation employed by Dakos et al. (2019). They may
also make for context dependent effects of functional diversity. I.e.
functional diversity may delay a tipping point in one context
(e.g.~ecosystem type and/or particular parameterisation of a specific
system) but delay it in another. \textbf{Based on our findings, (1) can
we make determinate predictions, and (2) what level of generality do our
preditions have?} {[}Possible text: The answers are (1) no, we cannot
make determinate predictions: we found sign varying effects of
functional trait diversity on the nature of ecosystem responses to
environmental change. And (2) our predictions have very limited
generality. This is because we did not aim for general predictions: we
investigate a single parameterisation of a single model, with many
explicit and implicit assumptions. This is why we need more
investigations, both theoretical and empirical, of effects of functional
trait diversity on environmental responses to environmental change.{]}

What are the priorities for follow-on modelling studies?

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Multifarious environmental driver changes.
\item
  More process based approach to variation in biodiversity: mutation,
  genetics,
\end{enumerate}

@@

\hypertarget{what-are-the-priorities-for-subsequent-research}{%
\subsection{What are the priorities for subsequent
research?}\label{what-are-the-priorities-for-subsequent-research}}

A research programme that integrates {[}text from ERC application{]}.
E.g. Global Change Microbiology -- with quote.

Then replicated across numerous ecosystems and taxa. That is, we imagine
a blue-print / template for proposals that aim to establish such an
integrative research programme. This template would contain mandatory
workpackages and experiments, and sufficient standardisation to allow
for rigorous and quantitative across programme meta-analyses

@@

\hypertarget{bibliography}{%
\subsection*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{subsection}{Bibliography}

@@

\leavevmode\hypertarget{ref-ceulemans2021}{}%
Ceulemans, Ruben, Laurie Anne Wojcik, and Ursula Gaedke. 2021.
{``Functional Diversity Alters the Effects of a Pulse Perturbation on
the Dynamics of Tritrophic Food Webs.''} \emph{bioRxiv}, March,
2021.03.22.436420. \url{https://doi.org/10.1101/2021.03.22.436420}.
